---
title: "grobert_d_ada_hw2"
author: "deving"
date: "3/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(psych)
library(cowplot)
```

#CHALLENGE 1
#read in the data
```
f <- "https://raw.githubusercontent.com/difiore/ada-2021-datasets/main/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)
```
#filter for movies from the 1920's- 1970's between 1 and 3 hrs long
```
d <- filter(d, startYear %in% 1920:1979)
d <- filter(d, runtimeMinutes >= 60 & runtimeMinutes <= 180)
```
#create "decade" variable
```
b<-c(-Inf, 1930, 1940, 1950, 1960, 1970, Inf)
names<-c("twenties", "thirties", "fourties", "fifties", "sixties", "seventies")
d$decade<-cut(d$startYear, breaks=b, labels=names)
```

#plot histogram of the distribution of runtime (in 10 minute bins) for each decade
```
p<-ggplot(data=d, aes(x=runtimeMinutes)) + 
  geom_histogram(color="black", fill="white", binwidth=10)+
    xlab("runtime (minutes)") +
    ylab("count")
p <- p + facet_wrap(~decade)
p
```

#"results" includes population mean and sd of runtime or each decade
```
results<-d %>%
  group_by(decade) %>%
  summarise_at(vars(runtimeMinutes), funs(mean(., na.rm=TRUE), sd(.,              na.rm=TRUE)))
```

#sample of 100 movies from each decade
```
s <- d %>% group_by(decade) %>% sample_n(size=100, replace=FALSE)
```
#sample mean and sd of runtime for each decade
```
s_results<-s %>%
  group_by(decade) %>%
  summarise_at(vars(runtimeMinutes), funs(mean(., na.rm=TRUE), sd(.,              na.rm=TRUE)))
```
#sample standard error of the mean (SEM) for each decade 
```
s_results$SEM<-s_results$sd/sqrt(100)
```

#compare sample estimates of sd, m, and SEM with population estimates of same and SE in the pop mean for n=100
```
compare<-inner_join(s_results,results, by = "decade")
compare<-compare%>%
  rename(
  sample_mean = mean.x,
  sample_sd = sd.x,
  sample_se = SEM,
  pop_mean = mean.y,
  pop_sd = sd.y,
  )
compare$pop_se<-compare$pop_sd/sqrt(100)
```

#sampling distribution: 1000 samples of 100 movies from each decade with mean and sd for each sample
```
n<-100
reps <- 1000

samp_dist <-
  do(reps) * d %>% 
  group_by(decade) %>% 
  sample_n(size=n, replace=FALSE)%>%
  summarise_at(vars(runtimeMinutes), funs(mean(., na.rm=TRUE), sd(.,              na.rm=TRUE)))
```

#mean and sd of sample means for each decade
```
samp_dist_results_mean<-samp_dist %>%
  group_by(decade) %>%
  summarise_at(vars(mean), funs(mean(., na.rm=TRUE)))
  
samp_dist_results_sd<-samp_dist %>%
  group_by(decade) %>%
  summarise_at(vars(mean), funs(sd(., na.rm=TRUE)))

samp_dist_results<-inner_join(samp_dist_results_mean, samp_dist_results_sd, by="decade")

samp_dist_results<-samp_dist_results%>%
  rename(
  dist_mean=mean.x,
  dist_sd=mean.y)
```  

#histograms of the sampling distributions for each decade
```
samp_dist_plot<-ggplot(data=samp_dist, aes(x=mean)) + 
  geom_histogram(color="black", fill="white", binwidth=1)+
    xlab("mean") +
    ylab("count")
samp_dist_plot <- samp_dist_plot + facet_wrap(~decade)
samp_dist_plot
```

#compare SE from each decade from initial sample, from known pop, and from samp distribution of means
```
compare_all<-inner_join(compare,samp_dist_results, by = "decade")

compare_all$dist_se<-compare_all$dist_sd/sqrt(1000)
```
#CHALLENGE 2
#probability that the bee biologist sees 9 bees or fewer
```
ppois(9, lambda=12)
```
#probability that the bee biologist sees 0 bees
```
ppois(0, lambda=12)
```
#probability that she sees exactly 5 bees
```
dpois(5, lambda=12)
```
#probability that she sees more than 18
```
ppois(18, lambda=12, lower=FALSE)
```
#plot fr the Poisson mass function for values 0 to 24
```
l <- 12
p1 <-
  plotDist(
    "pois",
    lambda = l,
    kind = "cdf",
    main = paste0("Cumulative Probability\nwith lambda=", l),
    xlab = "x",
    ylab = "Pr(X≤x)",
    type = "l"
  )
plot_grid(p1)
```

#simulate 1460 results
```
s<-rpois(1460, 12)
```

#plot simulation
```
s_plot<-histogram(s,
    type = "density",
    main = paste("Histogram of Outcomes of ", 1460, " Bee Sessions", sep = ""),
    xlab = "count",
    ylab = "Probability",
    xlim = 0:24
  )
```

#plots of the simulation and of the mass function are consistent, with about 50% cumulative probability reached at the mean (12) and about 100% reached at double the mean (24)

#CHALLENGE 3
#read in zombie data
```{r}
z <- "https://raw.githubusercontent.com/difiore/ada-2021-datasets/master/zombies.csv"
d <- read_csv(z, col_names = TRUE)
```
#mean for each quantitative variables
```{r}
m<-c((mean(d$height)), (mean(d$weight)),  (mean(d$age)), (mean(d$zombies_killed)), (mean(d$years_of_education)))
```
#create a function for pop sd
```{r}
pop.sd <- function(x){sd(x)*sqrt((length(x)-1)/length(x))}
```
#pop sd for quantitative variables
```{r}
sd <- c((pop.sd(d$height)), (pop.sd(d$weight)),  (pop.sd(d$age)), (pop.sd(d$zombies_killed)), (pop.sd(d$years_of_education)))
```
#boxplots for quantitative variables
```{r}
hp <- ggplot(data = d, aes(x = gender, y = height)) +
  geom_boxplot(na.rm = TRUE, outlier.shape = NA) 
hp
wp <- ggplot(data = d, aes(x = gender, y = weight)) +
  geom_boxplot(na.rm = TRUE, outlier.shape = NA) 
wp
ap <- ggplot(data = d, aes(x = gender, y = age)) +
  geom_boxplot(na.rm = TRUE, outlier.shape = NA) 
ap
zkp <- ggplot(data = d, aes(x = gender, y = zombies_killed)) +
  geom_boxplot(na.rm = TRUE, outlier.shape = NA) 
zkp
sp <- ggplot(data = d, aes(x = gender, y = years_of_education)) +
  geom_boxplot(na.rm = TRUE, outlier.shape = NA) 
sp
```

#scatterplot of height in relation to age, gender by point color
```{r}
sh <- ggplot(data = d, aes(
  x = age,
  y = height,
  color = factor(gender)
)) 
sh <- sh + geom_point(na.rm = TRUE)
sh
```
#scatterplots of weight in relation to age, gender by point color
```{r}
sw <- ggplot(data = d, aes(
  x = age,
  y = weight,
  color = factor(gender)
)) 
sw <- sw + geom_point(na.rm = TRUE)
sw
```
#both weight and height appear to be positively correlated to age, average values for both appear to be higher for males than females

#qq plot for height
```{r}
qqnorm(d$height)
qqline(d$height, col = "gray")
```

#histo for height
```{r}
histogram(d$height,
    type = "density",
 )
```
#height looks normal

#qq plot for weight
```{r}
qqnorm(d$weight)
qqline(d$weight, col = "gray")
```

#histo for weight
```{r}
histogram(d$weight,
    type = "density",
 )
```

#weight looks normal

#qq plot for age
```{r}
qqnorm(d$age)
qqline(d$age, col = "gray")
```
#histo for age
```{r}
histogram(d$age,
    type = "density",
 )
```
#age looks normal

#qq plot for zombies killed
```{r}
qqnorm(d$zombies_killed)
qqline(d$zombies_killed, col = "gray")
```
#histo for zombies killed
```{r}
histogram(d$zombies_killed,
    type = "density",
 )
```
#zombies killed NOT normal, looks like binomial

#qq plot for skoolin
```{r}
qqnorm(d$years_of_education)
qqline(d$years_of_education, col = "gray")
```
#histo for skoolin
```{r}
histogram(d$years_of_education,
    type = "density",
 )
```
#years of education NOT normal, looks like binomial

#generate subset of 50 survivors, calculate sd and mean
```{r}
s <- d %>% sample_n(size=50, replace=FALSE)
avg_h<- mean(s$height)
avg_w<- mean(s$weight)
avg_a<- mean(s$age)
avg_zk<- mean(s$zombies_killed)
avg_ye<- mean(s$years_of_education)
sd_h<- sd(s$height)
sd_w<- sd(s$weight)
sd_a<- sd(s$age)
sd_zk<- sd(s$zombies_killed)
sd_ye<- sd(s$years_of_education)
```
#create function for sample standard error of the mean
```{r}
sem <- function(x) sd(x)/sqrt(length(x))
```

#calculate sem for each quantitative variable
```{r}
sem_h<- sem(s$height)
sem_w<- sem(s$weight)
sem_a<- sem(s$age)
sem_zk<- sem(s$zombies_killed)
sem_ye<- sem(s$years_of_education)
```
#create a function to calculate 95% confidence intervals
```{r}
CI <- function(x, level = 0.95) {
  alpha <- 1 - level
  ci <- mean(x) + c(-1, 1) * qnorm(1 - (alpha / 2)) * sqrt(var(x) / length(x))
  return(ci)
}
```
#calculate 95% confidence intervals for each variable
```{r}
ci_h<- CI(s$height)
ci_w<- CI(s$weight)
ci_a<- CI(s$age)
ci_zk<- CI(s$zombies_killed)
ci_ye<- CI(s$years_of_education)
```
#draw 99 more samples of 50, calculate means
```{r}
n<-50
reps <- 99

samp_dist <-
  do(reps) * d %>% 
  sample_n(size=n, replace=FALSE)%>%
  summarise_at(vars(height, weight, age, zombies_killed, years_of_education), funs(mean(., na.rm=TRUE)))
samp_dist$.row<-NULL
samp_dist$.index<-NULL
```  
#combine with that initial sample
```{r}
initial_samp<-c(avg_h, avg_w, avg_a, avg_zk, avg_ye)
names(initial_samp)<-c("height", "weight", "age", "zombies_killed", "years_of_education")
samp_dist<-rbind(samp_dist, initial_samp)
view(samp_dist)  
```
#calculate mean and sd for sampling distribution
```{r}
samp_dist_results_mean<-samp_dist %>%
  summarise_at(vars(height, weight, age, zombies_killed, years_of_education), funs(mean(., na.rm=TRUE)))
  
samp_dist_results_sd<-samp_dist %>%
  summarise_at(vars(height, weight, age, zombies_killed, years_of_education), funs(sd(., na.rm=TRUE)))

samp_dist_results<-inner_join(samp_dist_results_mean, samp_dist_results_sd)
```
#construct 95% CI for each mean from the sampling distribution of sample means
```{r}
samp_dist_ci_h<- CI(samp_dist$height)
samp_dist_ci_w<- CI(samp_dist$weight)
samp_dist_ci_a<- CI(samp_dist$age)
samp_dist_ci_zk<- CI(samp_dist$zombies_killed)
samp_dist_ci_ye<- CI(samp_dist$years_of_education)
samp_dist_ci<-c(samp_dist_ci_h, samp_dist_ci_w, samp_dist_ci_a, samp_dist_ci_zk, samp_dist_ci_ye)
```
#how do these compare to the initial sample?
```{r}
view(samp_dist_ci)
initial_sample_ci<-c(ci_h, ci_w, ci_a, ci_zk, ci_ye)
view(initial_sample_ci)
```
# the confidence intervals form the sampling distribution are much tighter